---
title: "TITANIC"
author: "Meet"
date: "2023-03-28"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

# Working on data

#### Loading required library to work upon data in *R*

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(randomForest)
```

#### Importing data from the train model of *Titanic* dataset

```{r}
train_df <- read_csv("train.csv", show_col_types = FALSE)
test_df <- read_csv("test.csv", show_col_types = FALSE)
```

Now that we have imported the datasets, lets review them and observe some insights.

#### Review the dataset

```{r}
head(train_df)
glimpse(train_df)
str(train_df)
summary(train_df)
summary(test_df)
```

We can see in the summary that **177** missing values of Age is there in the *train_df* dataset but Age is an important factor in the survival on *Titanic*.

### Cleaning Data

#### Missing Values

Now lets make the **missing values** and the **space values** if any into **NA** to avoid any future confusions.\
We will create a copy of *train_df* and name it *train_df2*

```{r}
train_df2 <- train_df
train_df2[train_df2=="" | train_df2==" "] <- NA
```

Lets do the same for *test_df* dataset.

```{r}
test_df2 <- test_df
test_df2[test_df2=="" | test_df2==" "] <- NA
```

For future convenience, let's combine the dataset but the main problem is that the *test* dataset don't contain the column **Survived**.\
So we are going to bind the column with values **NA**.

```{r}
test_df2 <- cbind(test_df2, Survived = NA)
```

Now that we have added the column to the dataset, both the datasets contains same columns so now we will bind the rows to combine both data.

```{r}
alldata <- rbind(train_df2, test_df2)
```

Take a look at the summary of the combined data

```{r}
summary(alldata)
```

### Observing Data

Now we observed that the there is one missing value in *Fare* column. So let's observe the profile of the row to fill the data.

```{r}
alldata %>% 
  filter(is.na(Fare))
```

#### Assigning suitable value in Fare

```{r}
fare_df <- alldata %>% 
  filter(Embarked=="S", Sex=="male", Pclass==3, Age>=55)
head(fare_df)
```

From the data we can get idea about the **Median** of the data.\
Now let's check and verify the data that is there any variation in Fare prices in *Pclass=3*.

```{r warning=FALSE}
ggplot(data=alldata %>% filter(Pclass==3))+
  geom_histogram(mapping=aes(x=Fare))+
  labs(x="Fare Prices", y="No. of Passengers", title="Mapping of Fare Prices")
ggsave("Fare Prices.png")
```

It's around the **median** which is obtained by the filtered data.\
Now lets calculate the median and assign the value to **NA**.

```{r}
alldata$Fare[is.na(alldata$Fare)] <- median(fare_df$Fare, na.rm=T)

alldata %>% 
  filter(Age==60.5)
```

**NOTE**: *na.rm* is used to remove the missing values from the input vector.

#### Converting *Sex* column into numericals

Let's check the **NA** values in the *Sex* column if any.

```{r}
table(is.na(alldata$Sex))
```

There is no **NA** values.\
Now, assign the value **1** for *male* and **0** for *female*.

```{r}
alldata$Sex[alldata$Sex == "male"] <- 1
alldata$Sex[alldata$Sex == "female"] <- 0
head(alldata)
```

## Name

### Separating Title

Lets take a look at the sample of the 30 *Name* column to draw the conclusions about professional title they have.

```{r}
sample(alldata$Name, 30)
```

Let's separate the *Professional_title* data from *Name* column.

```{r}
alldata <- alldata %>% 
  separate(Name, into=c('name2', 'name3'), sep=', ')
```

```{r}
alldata <- alldata %>% 
  separate(name3, into=c('Professional_title', 'name4'), sep='. ')
```

```{r}
alldata <- alldata %>% 
  select(-name2,-name4)
```

### Mapping

Check for the **NA** if any,

```{r}
table(is.na(alldata$Professional_title))
```

As there is no **NA** so lets make a plot of total counts of all titles.

```{r}
ggplot(data=alldata) +
  geom_bar(mapping=aes(x=Professional_title), fill='blue') +
  labs(title="Professional Title") +
  theme(axis.text.x = element_text(angle=45))
ggsave("Professional Title.png")
```

Now, we observe the common title to have the idea about Survival for different title holders.

```{r message=FALSE, warning=FALSE}
ggplot(data = alldata %>% filter(Professional_title %in% c("Mr", "Miss", "Mrs", "Master"))) +
  geom_histogram(mapping=aes(x=Survived, fill=Professional_title))+
  facet_grid(~Professional_title)+
  labs(title="Mapping of common title", x="Survived=1")+
  theme(axis.text.x=element_text(angle=90))
ggsave("Professional Title Vs Survived.png")
```

### Converting the rare titles into popular ones

Let's observe the data of *male title* and draw some insights about them.

```{r}
alldata %>% 
  filter(Professional_title %in% c("Capt", "Col", "Don", "Dr", "Jonkheer", "Major", "Rev", "Sir"))
```

Let's rename the titles like *Capt*, *Col*, *Don*, *Dr*, *Jonkheer*, *Major*, *Rev*, *Sir* into *Mr* for making our life easy as there is no important insight and No. of Survived are equivalent to *Mr*....

```{r}
alldata$Professional_title[alldata$Professional_title %in% c("Capt", "Col", "Don", "Dr", "Jonkheer", "Major", "Rev", "Sir")] <- "Mr"
```

Do the same process with *female title*.

```{r}
alldata %>% 
  filter(Professional_title %in% c("Dona", "Lady", "Mlle", "Mme", "th"))
```

Let's rename the titles like *Dona*, *Lady*, *Mlle*, *Mme*, *th* into *Mrs* and *Ms* into *Miss* as Ms is basically the short of Miss.

```{r}
alldata$Professional_title[alldata$Professional_title %in% c("Dona", "Lady", "Mlle", "Mme", "th")] <- "Mrs"
```

```{r}
alldata$Professional_title[alldata$Professional_title == "Ms"] <- "Miss"
```

```{r}
ggplot(data=alldata) +
  geom_bar(mapping=aes(x=Professional_title)) +
  labs(title="Professional Title (Only Populars)") +
  theme(axis.text.x = element_text(angle=45))
ggsave("Professional Title (Only Populars).png")
```

## Embarked

First of all see the **NA** value if any in the *Embarked* column

```{r}
table(is.na(alldata$Embarked))
```

### Observe the data

Review the both rows and draw the conclusion to fill the missing values.

```{r}
alldata %>% 
  filter(is.na(Embarked))
```

We can see that both have *Pclass*=1, same *Ticket* number, *Fare*=80, same *Cabin* and both *female*.

```{r}
ggplot(alldata %>% filter(Pclass==1)) +
  geom_boxplot(mapping=aes(x=Embarked, y=Fare, fill=Sex)) +
  labs(title="Mapping of data taking Pclass=1")
ggsave("Pclass.png")
```

With the help of plot we can conclude that missing value is the **C**.

### Filling the missing value

```{r}
alldata$Embarked[is.na(alldata$Embarked)] <- "C"
```

## Age

Let's see the **NA** value if any in the *Age* column.

```{r}
table(is.na(alldata$Age))
```

Let' draw a plot to have a better idea about *age* and *title* as they can have some relationship to fill our missing.

```{r}
ggplot(alldata) +
  geom_boxplot(mapping=aes(x=Professional_title, y=Age)) +
  labs(title = "Median of the different titles")
ggsave("Median of Titles.png")
```

Let's take the **median** of the respective *Professional_title* and fill the missing value respectively.

### Filling Master title

```{r}
master_df <- alldata %>% 
  filter(Professional_title=="Master")
master_df$Age[is.na(master_df$Age)] <- median(master_df$Age, na.rm=T)
```

### Filling Miss title

```{r}
miss_df <- alldata %>% 
  filter(Professional_title=="Miss")
miss_df$Age[is.na(miss_df$Age)] <- median(miss_df$Age, na.rm=T)
```

### Filling Mr title

```{r}
mr_df <- alldata %>% 
  filter(Professional_title=="Mr")
mr_df$Age[is.na(mr_df$Age)] <- median(mr_df$Age, na.rm=T)
```

### Filling Mrs title

```{r}
mrs_df <- alldata %>% 
  filter(Professional_title=="Mrs")
mrs_df$Age[is.na(mrs_df$Age)] <- median(mrs_df$Age, na.rm=T)
```

### Binding

Now, merge the data into again *alldata* dataset.

```{r}
alldata <- rbind(master_df, miss_df, mr_df, mrs_df)
alldata <- alldata %>% 
  arrange(PassengerId)
```

## Ticket

Check the **NA** if any,

```{r}
table(is.na(alldata$Ticket))
```

Take the sample and observe the data of *ticket* column.

```{r}
sample(alldata$Ticket, 30)
```

Can't have any relevance to the survival of the passengers, so decided to remove the column

```{r}
alldata <- alldata %>% 
  select(-Ticket)
```

## Pclass

Check the **NA** if any,

```{r}
table(is.na(alldata$Pclass))
```

## SibSp

Check the **NA** if any,

```{r}
table(is.na(alldata$SibSp))
```

## Parch

Check the **NA** if any,

```{r}
table(is.na(alldata$Parch))
```

## Cabin

Check the **NA** if any,

```{r}
table(is.na(alldata$Cabin))
```

As there are many empty cells in the *Cabin* column so best possible solution is to drop the column only.

```{r}
alldata <- alldata %>% 
  select(-Cabin)
```

# Appling Model into our cleaned data

### Let's apply a randomForest

Honestly didn't know the shit about this models right now :(\
but let's apply to our data.

```{r}
i <- is.na(alldata$Survived)
myforest <- randomForest(data=alldata[!i,], Survived ~ Pclass + Age + SibSp + Parch + Fare + Embarked + Sex + Professional_title,
                             ntree=10000, sampsize = 400, mtry=4)
```

Now let's look the **model** works decent or not...

```{r}
table(round(predict(myforest, newdata=alldata[!i,])) == alldata[!i,]$Survived)
```

Add the predictions into our data for submissions

```{r}
alldata$forestpred <- round(predict(myforest, newdata=alldata))
rm(myforest)
```

Export the **.csv** file

```{r}
write.csv(alldata %>% 
  filter(PassengerId %in% c(892:1310)) %>% 
  select(PassengerId, forestpred) %>% 
  rename(Survived = forestpred), "submission.csv")
```

Now deleted the No. column in the Excel manually and submitted the data........**0.77272**\
Great for my first Project.

Thank you
